\documentclass[12pt]{amsart}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{nopageno}
\usepackage{amssymb}
\usepackage{epstopdf}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\title{18.06 Recitation Feedback}
%\author{The Author}
%\date{}                                           % Activate to display a given date or no date

\begin{document}

\maketitle
%\section{}
%\subsection{}
\thispagestyle{empty}
\begin{enumerate}

\item On a scale of $0$ to $10$, how much to you agree with the following statements.
\vspace{10pt}
\begin{itemize}
\item[\underline{\phantom{aaaaa}}] I feel comfortable asking questions in recitation.
\vspace{5pt}
\item[\underline{\phantom{aaaaa}}] I feel comfortable asking questions in office hours.
\vspace{5pt}
\item[\underline{\phantom{aaaaa}}] I feel comfortable asking questions via email or piazza.
\end{itemize}

\vspace{10pt}

\item (Mark all that apply) Recitation would be more useful to me if:
\begin{itemize}
\item[$\Box$] Less problems were solved.
\vspace{5pt}
\item[$\Box$] More problems were solved.
\vspace{5pt}
\item[$\Box$] I had more time to work on problems on my own before they are solved.
\vspace{5pt}
\item[$\Box$] I had less time to work on problems on my own before they are solved.
\vspace{5pt}
\item[$\Box$] Problems were solved more slowly.
\vspace{5pt}
\item[$\Box$] Problems were solved more quickly.
\vspace{5pt}
\item[$\Box$] There were less audience participation.
\vspace{5pt}
\item[$\Box$] There were more audience participation.
\vspace{5pt}
\item[$\Box$] More time were devoted to material/definition review.
\vspace{5pt}
\item[$\Box$] Less time were devoted to material/definition review.
\end{itemize}

\vspace{10pt}

\item On a scale of $0$ to $10$, how much to you agree with the following statements.
\vspace{10pt}
\begin{itemize}
\item[\underline{\phantom{aaaaa}}] I don't come to (Isabel's) office hours because of the time.
\vspace{5pt}
\item[\underline{\phantom{aaaaa}}] I don't come to (Isabel's) office hours because I know how to do the pset.
\vspace{5pt}
\item[\underline{\phantom{aaaaa}}] I would come to (Isabel's) office hours (more frequently) if we talked about class material I was confused about (instead of pset questions).
\vspace{5pt}
\item[\underline{\phantom{aaaaa}}] I would come to additional office hours to discuss class material I am confused about.
\end{itemize}
\vspace{10pt}
\item Comment on what works in recitation, and what doesn't.  How can Isabel be a maximally helpful TA?


\end{enumerate}
\end{document}  