\documentclass[11pt]{article}
\usepackage[hmargin=40pt,vmargin=40pt]{geometry}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{enumerate}
\pagenumbering{gobble} 
\newcommand{\diff}{\,\mathrm{d}}
\renewcommand*{\vec}[1]{\mathbf{#1}}

\title{18.06 R08 - Recitation 1}
\author{Sam Turton}
\date{February 12, 2019}                                      
\begin{document}
\maketitle


\rule{\textwidth}{1pt}
\section{Important information}
Welcome to recitations for 18.06! 
\begin{itemize}
\item Recitation leader: Sam Turton
\item Office: 2-333C 
\item Email: seturton@mit.edu
\item Office hours: Tuesdays, 3:30-5:30pm, 2-361
\end{itemize}

\noindent We are \emph{not} using stellar for this class (apart from recitation sign-up):
\begin{itemize}
\item All course materials, lecture summaries, psets, pset solutions, \emph{etc.} will be posted on the course website (https://mitmath.github.io/1806/).
\item HW should be submitted on gradescope (https://www.gradescope.com/courses/40389).
\end{itemize}
 
\section{Lecture summary}
\begin{enumerate}
\item In essence, a \textbf{vector space} is a set $V$ of vectors for which any linear combination of vectors in $V$ is also in $V$. Examples:
\begin{itemize}
\item vectors in $\mathbb{R}^n$.
\item matrices in $\mathbb{R}^{m\times n}$
\item real valued functions $f(x)$
\end{itemize}
\item A \textbf{vector subspace}, $X$, is a set of vectors that satisfies the following requirement: \emph{If $v$ and $w$ are vectors in $X$, and $c$ is any scalar, then}
\begin{itemize}
\item \emph{$v+w$ is in $X$}
\item \emph{$cv$ is in $X$}
\end{itemize}
\item A \textbf{linear system} is an equation of the form $Ax=b$, where $A$ is an $m\times n$ matrix, $x$ is a vector in $\mathbb{R}^n$, and $b$ is a vector in $\mathbb{R}^m$. Much of the purpose of linear algebra is understanding when this equation has solutions, and what these solutions look like. 
\item If $m=n$, then it is possible that $A$ will have an \textbf{inverse} $A^{-1}$, so that $A^{-1}A=A^{-1}A = I$. \textbf{This does not always exist}. If it does exist, then $x=A^{-1}b$. But this is usually still not a good way to find a solution.... 
\item If $Q$ is a square matrix with \emph{orthonormal} columns, then $QQ^T=Q^TQ = I$ and we say that $Q$ is an \textbf{orthogonal} matrix. In particular, $Q^{-1}=Q^T$. 
\end{enumerate}

\newpage

\section{Problems}

\noindent \textbf{Problem 1. (Vector spaces)}\\
$\mathbb{R}^3$ is the vector space of all vectors of the form $b = (b_1 , b_2 , b_3 )$. Which of the following are subspaces of $\mathbb{R}^3$?
\begin{enumerate}[(a)]
\item The set of all vectors $(b_1, b_2, b_3)$ where $b_1 = b_2$. 
\item The set of all vectors with $b_1 = 1$.
\item The set of all vectors with $b_1b_2b_3 = 0$.
\item All linear combinations of $v = (1, 4, 0)$ and $w = (2, 2, 2)$.
\item All vectors that satisfy $b_1 + b_2 + b_3 = 0$. 
\item All vectors with $b_1 \leq b_2 \leq b_3$.
\end{enumerate}

\vskip 100pt

\noindent \textbf{Problem 2. (Linear systems and block matrices)}\\
\begin{enumerate}[(a)]
\item Consider the simultaneous equations
\begin{align*}
ax + by &= e\\
cx + dy &= f,
\end{align*}
where $a,b,c,d,e$ and $f$ are real numbers and $x,y$ are scalar unknowns. Write these equations as a single matrix equation $Az = g$. What are $A, z$ and $g$? What are their dimensions?
\item If you have two coupled systems of equations
\begin{align*}
Bx + Cy &= c\\
Dx + Ey &= d,
\end{align*}
where $B, C, D$ and $E$ are $3 \times 3$ matrices and $x, y, c$ and $d$ are 3-component vectors, can you write this
as a single system of equations,
$$Az = b,$$
where $z = \begin{pmatrix} x \\ y \end{pmatrix}$ is the 6-component vector of $x$ on top of $y$? What are $A$ and $b$? What are their dimensions?
\end{enumerate}

\newpage 

\noindent \textbf{Problem 3. (LU factorization)}\\
\begin{enumerate}[(a)]
\item Consider the linear system $Ax = b$, where
$$A = \begin{pmatrix} 1 & -1 & 2 \\ 0 & 1 & 3 \\ 0 & 0 & 1 \end{pmatrix}, \;\;\; b = \begin{pmatrix} -1 \\ 3 \\ 2 \end{pmatrix}, \;\;\; x = \begin{pmatrix} x_1 \\ x_2 \\ x_3 \end{pmatrix}.$$
Write the system as three simultaneous equations for the unknowns $x_1, x_2$ and $x_3$. Solve these equations explicitly in order to find a solution to the linear system.
\item Consider the two $3\times 3$ matrices $L$ and $U$, where
$$L = \begin{pmatrix} 1 & 0 & 0 \\ 0 & 1 & 0 \\ -1 & 1 & 1 \end{pmatrix}, \;\;\; U = \begin{pmatrix} 1 & -1 & -2 \\ 0 & 1 & 1 \\0 & 0 & 2 \end{pmatrix}.$$
Compute $B = LU$.
\item Consider the linear system $Bx = b$, where $B=LU$ is what you computed in the previous part, $b$ is the same vector as in part (a), and $x$ is an unknown vector. Find the solution to this linear system using the following steps:
\begin{itemize}
\item Explicitly solve the  linear system $Ly = b$ to find $y$.
\item Using the $y$ you found in the previous step, explicitly solve $Ux = y$ for $x$.
\item Then this $x$ solves the linear system $Bx = b$. Check this! Why does this work?\footnote{Hint: Note that $Bx = b$ is the same as $LUx = L(Ux) = b$.}
\end{itemize}
\end{enumerate}

\newpage 

\noindent \textbf{Problem 4. (More matrix factorization)}\\
Under what conditions on $a,b,c$ and $d$ can a $2\times 2$ matrix $A=\begin{pmatrix} a & b \\ c & d\end{pmatrix}$ be written as $A=X^TX$, where $X=\begin{pmatrix} p & 0 \\ q & r \end{pmatrix}$? What are $p,q$ and $r$ in terms of $a,b,c$ and $d$?




\end{document}  